---
title: "Coronavirus COVID-19 Brazil Cases"
output: 
  flexdashboard::flex_dashboard:
    css: styles.css
    vertical_layout: scroll
    orientation: rows
    navbar:
    - { title: "PROCC/Fiocruz", href: "http://www.procc.fiocruz.br/" }
editor_options: 
  chunk_output_type: console
---

```{r, echo=FALSE, results='hide'}
library(curl)
library(DT)
library(flexdashboard)
library(ggplot2)
library(ggrepel)
library(leaflet)
library(leafpop)
library(plotly)
library(tidyr)
library(tidyverse)
library(htmltools)
library(sp)
library(rgdal)
library(data.table)
library(shiny)
```


```{r, echo=FALSE, results='hide', cache=FALSE}

# Reading data

dconfirmed <- read.csv("https://raw.githubusercontent.com/wcota/covid19br/master/cases-brazil-cities-time.csv",sep=",",encoding = "UTF-8") %>% filter(!state == "TOTAL") %>% mutate(status = "Confirmed") 
# %>% mutate(city = substr(city,1,nchar(as.character(city))-3))

# ddeaths <- read.csv("")

# drecovered <- read.csv("")

# Remove rows where Value is empty
if(length(which(dconfirmed$totalCases == ""))>0){
  dconfirmed <- dconfirmed[-which(dconfirmed$totalCases == ""), ]
}
# if(length(which(ddeaths$Value == ""))>0){
# ddeaths <- ddeaths[-which(ddeaths$Value == ""), ]
# }
# if(length(which(drecovered$Value == ""))>0){
# drecovered <- drecovered[-which(drecovered$Value == ""), ]
# }

pop2019 <- read.csv("https://raw.githubusercontent.com/lbianchi/Infodengue/master/EstPopBrasil2019TCU.csv",sep=";",dec=",",encoding = "UTF-8") %>% mutate(MUNICIPIO = paste0(MUNICIPIO,"/",UF),
         POPULACAO = as.numeric(POPULACAO),
         LONG = as.numeric(LONG),
         LAT = as.numeric(LAT))
colnames(pop2019) <- c("state","cod.state","cod.city","city","pop","geocode","long","lat")
dados <- dconfirmed %>% left_join(pop2019, by = "city") %>% mutate(state = state.x,Deaths = 0, Recovered = 0) %>% select(-c(state.x,state.y)) %>% group_by(city,date) %>% 
  mutate(
    casesper100k = totalCases/pop*100000,
    newCasesper100k = newCases/pop*100000
  ) %>% ungroup() %>% 
  mutate(city = as.character(city)) %>% data.frame()
```


```{r, echo=FALSE, results='hide'}

# ggplot time

# id locations

# selecting cities with 3 or more confirmed cases
vecid <- dados %>% group_by(state,city) %>% summarise(n=n()) %>% 
  filter(n>2) %>% ungroup() %>% select(city) 

glist <- NULL
for(i in 1:nrow(vecid)){
  # d1 <- rbind(dconfirmed, ddeaths,drecovered)
  # replace dconfirmed by d1 when ddeaths or drecovered become available
  d1 <- dados[which(dados$city == vecid$city[i]), ]
  
  glist[[i]] <- ggplot() + 
    geom_point(data = d1, aes(x = date, y = totalCases),stat = "identity",size = 5, colour = "#e02bbc") +
    geom_line(data = d1, aes(x = date, y = totalCases,group = 1),size = 2, colour = "#e02bbc") +
    geom_bar(data = d1,aes(x = date, y = newCases), stat = "identity", position = "dodge", fill = "#612be0") +
    ggtitle(paste(as.character(vecid$city[i]), "\n \nConfirmados:", sum(d1$totalCases))) +
    theme_minimal() +
    theme(plot.title = element_text(size = 14,face = "bold"),
          text = element_text(size = 12),
          axis.title = element_blank(),
          axis.text.x = element_blank(),
          axis.text.y = element_blank(),
          panel.grid.minor = element_blank(),
          legend.position = "bottom",
          legend.title = element_blank(),
          legend.text = element_blank(),
          plot.margin = margin(0, 0, 0, 0, "cm"))
}
```


```{r, echo=FALSE, results='hide'}

# leaflet map dconfirmed, ddeaths, drecovered

dados$color <- "b"
dados$color[which(dados$totalCases > 100)] <- "a"
pal <- colorFactor(palette = c("red", "orange"), domain = dados$color)

l <- leaflet(dados) %>%
  addTiles() %>%
  addProviderTiles(providers$CartoDB.Positron, group = "Light") %>%
  addProviderTiles(providers$CartoDB.DarkMatter, group = "Dark") %>%
  addProviderTiles(providers$HERE.satelliteDay, group = "Satellite") %>%
   addCircles(lng = ~long, lat = ~lat, radius = ~ totalCases*100, group = "pt", col = ~ pal(color), popup = ~paste0("<b>",city,":</b><br>",
"<tr><td>Confirmed:</td><td align=`right`>", dados$totalCases, "</td></tr><br>",
"<tr><td>New cases:</td><td align=`right`>", dados$newCases, "</td></tr><br>",
"<tr><td>Deaths:</td><td align=`right`>", dados$Deaths, "</td></tr><br>",
"<tr><td>Recovered:</td><td align=`right`>", dados$Recovered, "</td></tr>"
)) %>%
  #addPopupGraphs(glist, group = "pt", width = 400, height = 300) %>%
  setView(lng = -56.0949, lat = -15.5989, zoom = 4) %>% 
  addLayersControl(
    baseGroups = c("Dark (default)", "Light", "Satellite"),
    overlayGroups = c("Confirmed","New cases", "Deaths", "Recovered"),
    options = layersControlOptions(collapsed = TRUE)
  )%>%
  hideGroup("Confirmed") %>%
  hideGroup("New cases") %>%
  hideGroup("Deaths") %>%
  hideGroup("Recovered") %>%
  addEasyButton(easyButton(
    icon    = "glyphicon glyphicon-globe", title = "Reset zoom",
    onClick = JS("function(btn, map){ map.setView([20, 0], 2); }"))) %>%
  addEasyButton(easyButton(
    icon    = "glyphicon glyphicon-map-marker", title = "Locate Me",
    onClick = JS("function(btn, map){ map.locate({setView: true, maxZoom: 6}); }")))
```


```{r}

# DT Table
lastday <- sort(dados$date)[length(dados$date)]
t <- dados %>% select(city,newCases,totalCases,Deaths,Recovered) %>% group_by(city) %>% summarise(
  newCases = sum(newCases),
  totalCases = sum(totalCases),
  Deaths = sum(Deaths),
  Recovered = sum(Recovered)
) %>% arrange(desc(totalCases)) %>% data.frame()
colnames(t) <- c("City", "New Cases", "Confirmed","Deaths", "Recovered")
# t <- merge(dconfirmed[which(dconfirmed$date == lastday), ], ddeaths[which(ddeaths$date == lastday), ], by = "areaid")
# t <- merge(t, drecoverd[which(drecoverd$date == lastday), ], by = "areaid")

# t$confirmed <- t$cases.x
# t$deaths <- t$cases.y
# t$recovered <- t$cases

# t <- t[rev(order(t$confirmed)), ]
# t <- t[, c( "areaid", "confirmed", "deaths", "recovered")]
# rownames(t) <- NULL
# colnames(t) <- c("Country", "Confirmed", "Deaths", "Recovered")
# rownames(t) <- NULL
```


```{r}

# Time trends for each country with number of cases by number of days since 100th case
# https://cran.rstudio.com/web/packages/ggrepel/vignettes/ggrepel.html
# https://mran.microsoft.com/snapshot/2017-08-20/web/packages/ggrepel/vignettes/ggrepel.html

fnCompareAreaid <- function(d1, vblexcompare, vbleycompare, yint, vecareaid, titulo){
  
  # Time trends
  
  # Select data for some countries
  d1 <- d1[which(d1$city %in% veccity), ]
  
  g <- ggplot(d1[which(d1[, vblexcompare] >= 0), ], aes_string(x = vblexcompare, y = vbleycompare, group = "city", color = "city")) + 
    geom_line(size = 0.8) +
    geom_point(pch = 21, size = 1) +
    geom_hline(yintercept = yint) +
    geom_vline(xintercept = 0) +
    #scale_y_log10() +
    scale_y_log10(expand = expand_scale(add = c(0, 0.1)), breaks = c(10, 50, 100, 200, 500, 1000, 2000, 5000, 10000, 20000, 50000)) +
    #scale_y_continuous(expand = expand_scale(add = c(0,100))) +
    scale_x_continuous(expand = expand_scale(add = c(0, 1))) + # limit = c(0, 40)
    theme_minimal() +
    theme(panel.grid.minor = element_blank(), legend.position = "none", plot.margin = margin(3, 15, 3, 3, "mm")) +
    #geom_shadowtext(aes(label = paste0(" ", areaname)), hjust = 0, vjust = 0, data = . %>% group_by(city) %>%
    #                top_n(1, dayssince100), bg.color = "white") +
    labs(x = "", y = "", title = titulo) +
    coord_cartesian(clip = "off") +
    geom_text_repel(data = do.call(rbind, lapply(split(d1, as.factor(d1$city)), function(x){return(x[which.max(x[, vbleycompare]),])})),
                    aes(label = paste(" ", areaname)), size = 4, nudge_x = 1, segment.color = NA) # xlim = c(18, NA) 
  return(g) 
}
```


```{r}
fnTimeTrendCases <- function(d1, vblexcompare, vbleycompare, ai, titulo){
  # d1 = dados
  # vblexcompare = "date"
  # vbleycompare = "totalCases"
  # ai = as.character(t$City[1])
  # titulo = ""
  # d1 <- rbind(dconfirmed, ddeaths, drecovered)
  d1 <- d1[which(d1$city == ai), ]
  
  if(nrow(d1)<1){
    return(print("There is no Covid-19 case for this city."))
  }else{
    numconfirmed <- d1[which(d1$date == sort(d1$date)[length(d1$date)]), "totalCases"]
    # numdeaths    <- d1[which(d1$status == "Deaths"    & d1$date == sort(d1$date)[length(d1$date)]), "cases"]
    # numrecovered <- d1[which(d1$status == "Recovered" & d1$date == sort(d1$date)[length(d1$date)]), "cases"]
    
    # vbleycompare is "date" vbleycompare is "newCases" or "totalCases"
    if(vbleycompare=="totalCases"){
    g <- ggplot() + 
  geom_point(data = d1, aes(x = date, y = totalCases),stat = "identity",size = 1, colour = "#e02bbc") +
  geom_line(data = d1, aes(x = date, y = totalCases,group = 1),size = 1, colour = "#e02bbc") +
  geom_bar(data = d1,aes(x = date, y = newCases), stat = "identity", position = "dodge", fill = "#612be0") +
  ggtitle(paste(as.character(ai), "\n \nConfirmated:", sum(d1$totalCases))) +
  theme_minimal() +
  theme(plot.title = element_text(size = 14,face = "bold"),
        text = element_text(size = 12),
        axis.title = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        panel.grid.minor = element_blank(),
        legend.position = "bottom",
        legend.title = element_blank(),
        legend.text = element_blank(),
        plot.margin = margin(0, 0, 0, 0, "cm"))
    }else{
      if(vbleycompare=="newCases"){
        g <- ggplot() + 
  geom_bar(data = d1,aes(x = date, y = newCases), stat = "identity", position = "dodge", fill = "#612be0") +
  ggtitle(paste(as.character(ai), "\n \nNew cases:", sum(d1$newCases))) +
  theme_minimal() +
  theme(plot.title = element_text(size = 14,face = "bold"),
        text = element_text(size = 12),
        axis.title = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        panel.grid.minor = element_blank(),
        legend.position = "bottom",
        legend.title = element_blank(),
        legend.text = element_blank(),
        plot.margin = margin(0, 0, 0, 0, "cm"))
        }else{
        return(print("Argument invalid. Choose newCases or totalCases"))
      }
    }
    g <- ggplotly(g)
    return(g)
    #g <- ggplotly(g) %>% layout(legend = list(orientation = 'h'),
    #title = list(text = paste0(paste(unique(d1[d1$city == ai, "areaname"])),
    #             '<br>', '<sup>', paste("Confirmed:", numconfirmed,  " Deaths:", numdeaths,  " Recovered:", numrecovered), '</sup>')))
  }
}
```


Map
=====================================  


Row
-----------------------------------------------------------------------

### Confirmed

```{r}
valueBox(value = sum(dados$newCases),
         color = "#f5c147",
         icon = "glyphicon-ok")
```

### Deaths

```{r}
valueBox(value = sum(dados$Deaths),
         color = "#f5c147",
         icon = "glyphicon-heart")
```

### New Cases


```{r}
#usar o dconfirmed pq o dados remove
valueBox(value = sum(dados$newCases[which(dados$date==lastday)]),
         color = "#f5c147",
         icon = "glyphicon-ok-sign")
```


Row
-----------------------------------------------------------------------

### Click circles to see time trends {data-width=600}


```{r}
l
```


### Search and order {data-width=400}

```{r}
datatable(t, options = list(pageLength = 20))
```


States
===================================== 

Row
-----------------------------------------------------------------------


### Accumulated cases

```{r}
fnTimeTrendCases(dados,
                 vblexcompare = "date",
                 vbleycompare = "totalCases",
                 ai = as.character(t$City[1]),
                 titulo = "")
```

### Accumulated cases

```{r}
fnTimeTrendCases(dados,
                 vblexcompare = "date",
                 vbleycompare = "totalCases",
                 ai = as.character(t$City[2]),
                 titulo = "")
```

### Accumulated cases

```{r}
fnTimeTrendCases(dados,
                 vblexcompare = "date",
                 vbleycompare = "totalCases",
                 ai = as.character(t$City[3]),
                 titulo = "")
```

### Accumulated cases

```{r}
fnTimeTrendCases(dados,
                 vblexcompare = "date",
                 vbleycompare = "totalCases",
                 ai = as.character(t$City[4]),
                 titulo = "")
```

### Accumulated cases

```{r}
fnTimeTrendCases(dados,
                 vblexcompare = "date",
                 vbleycompare = "totalCases",
                 ai = as.character(t$City[5]),
                 titulo = "")
```


Row
-----------------------------------------------------------------------

### New cases

```{r}
fnTimeTrendCases(dados,
                 vblexcompare = "date",
                 vbleycompare = "newCases",
                 ai = as.character(t$City[1]),
                 titulo = "")
```

### New cases

```{r}
fnTimeTrendCases(dados,
                 vblexcompare = "date",
                 vbleycompare = "newCases",
                 ai = as.character(t$City[2]),
                 titulo = "")
```

### New cases

```{r}
fnTimeTrendCases(dados,
                 vblexcompare = "date",
                 vbleycompare = "newCases",
                 ai = as.character(t$City[3]),
                 titulo = "")
```

### New cases

```{r}
fnTimeTrendCases(dados,
                 vblexcompare = "date",
                 vbleycompare = "newCases",
                 ai = as.character(t$City[4]),
                 titulo = "")
```

### New cases

```{r}
fnTimeTrendCases(dados,
                 vblexcompare = "date",
                 vbleycompare = "newCases",
                 ai = as.character(t$City[5]),
                 titulo = "")
```



```{r}
#Comparison regions
#===================================== 
#Row
#-----------------------------------------------------------------------
### Number of confirmed cases by number of days since 100th case
#fnCompareAreaid(d, "dayssince100", "cases", 100, diso3codesFew$areaid, paste0("Confirmed cases by number of days since ", 100 ,"th case"))
```


```{r}
### Number of deaths by number of days since 10th case
#fnCompareAreaid(dd, "dayssince10", "cases", 10, diso3codesFew$areaid, paste0("Deaths by number of days since ", 10 ,"th death"))
```

Cases adjusted for population
===================================== 

```{r}
fnBarPlot <- function(d1, vbleplot, titulo){
  d1 = d1 %>% arrange(city,desc(date)) %>% 
    mutate(temp = 0)
  
  for(i in 2:nrow(d1)){
    if(d1$city[i]==d1$city[i-1]){
      d1$temp[i] <- 1
    }else{
      d1$temp[i] <- 0
    }
  }
  d1 <- d1 %>% filter(temp==0) %>% select(-temp) %>% arrange(desc(totalCases))
  d1 <- d1[1:10,] %>% data.frame()
  
  g <- ggplot(data = d1, aes_string(x = "city", y = vbleplot,
                                    fill = vbleplot)) + 
    geom_bar(stat = "identity", fill = "#612be0") +
    theme_minimal() + labs(title = titulo,y = "Cases per 100k habs.") +
    theme(axis.title.x = element_text(size = 10),
          axis.text.x = element_blank(),
          axis.title.y=element_blank(),
          legend.position = "none",
          legend.title = element_blank(),
          legend.text=element_text(size = 10),
          plot.margin = margin(0, 0, 0, 0, "cm")
      )  + coord_flip(clip = "off")
  
  
  return(g)
}
```

Row
-----------------------------------------------------------------------

### Confirmed cases per 100.000 habs (number of cases)

```{r, out.width='100%'}
fnBarPlot(dados, "casesper100k", "Confirmed cases per 100.000 habs. (number of cases)")
```


### Deaths per 100.000 habs (number of deaths)

```{r}
# fnBarPlot(d1[which(d1$status == "Deaths"), ], "casesper1m", "Deaths per one million population (number of deaths)")
```


Information  Last updated at `r sort(dados$date)[length(dados$date)]` 
===================================== 

**Author**

Leo Bastos <br>
Twitter: @leosbastos  
Website: https://lsbastos.github.io

Lucas Bianchi  
Twitter: @lucasmbianchi_  
Website: http://www.lucasbianchi.com/

Paula Moraga  
Twitter: @Paula_Moraga_  
Website: http://www.paulamoraga.com/

**Covid-19 Cases in Brazil**

https://github.com/wcota/covid19br/blob/master/cases-brazil-cities-time.csv


**Estimated population in 2019 in Brazil (TCU)**

ftp://ftp.ibge.gov.br/Estimativas_de_Populacao/Estimativas_2019/


**The R Project for Statistical Computing**

https://www.r-project.org/



**R packages**

curl: https://cran.r-project.org/web/packages/curl/index.html  
DT: https://rstudio.github.io/DT/  
flexdashboard: https://rmarkdown.rstudio.com/flexdashboard/    
ggplot2: https://ggplot2.tidyverse.org/  
leaflet: https://rstudio.github.io/leaflet/  
leafpop: https://github.com/r-spatial/leafpop

